---
title: "Takeaways from rstudio::conf2019"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

*From the webinars, because I did not attend* :sob:

# R Markdown

## Advanced R Markdown Workshop, Yihui Xie and Alison Hill

[Website](https://arm.rbind.io/)

Topics I don't cover in this summary: 

- [officer](https://rpodcast.github.io/officer-advrmarkdown)
- [flexdashboard](https://arm.rbind.io/slides/flexdashboard.html)
- [learnr](https://arm.rbind.io/slides/learnr.pdf)

### xaringan

*[Slides](https://arm.rbind.io/slides/xaringan.html)*

In the code for slides:

| Term or command | What does it do? |
|-----------------|------------------|
| `name: mynameis` | Name of slide, useful for cross-referencing.|
| `class: center, right` | Sets alignment of the entire slide. |
| `background-image: url(url/or/path)` | Add background image. |
| `layout: true` | Will apply layout to all slides. Useful for logos. |
| `???` | Presenter notes. |
| `.left[]` `.right[]` | Align some text only. |
| `.pull-left[]` `.pull-right[]` | Two-columns. |

Inside the code chucks:

| Term or command | What does it do? |
|-----------------|------------------|
| `#<<` | Highlight a line of code input. |
| `knitr::kable(format = "html")` | Print a table. |

In the chunk options:

| Term or command | What does it do? |
|-----------------|------------------|
| `highlight.output = c(1, 3)` | Highlight lines 1 and 3. |
| `highlight.output = 1:3` | Highlight lines 1 to 3. |
| `highlight.output = TRUE` | Highlight all lines of the output. |
| `r ref.label='another-chunk'` | Prints and evaluates a previous code chunk.|


YAML options for `output:xaringan::moon_reader:`:

| Example | What does it do? |
|-----------------|------------------|
| `nature:highlightStyle:github` | Define highlight style option. |
| `nature:highlightLines:true` | Enable/disable line highlighting. |
| `seal:false` | First slide will not use the information from the YAML header.|
| `css:[default, rladies, rladies-fonts]` | Use R-Ladies theme. |

Global options:

| Term or command | What does it do? |
|-----------------|------------------|
| `options(knitr.table.format = "html")` | Knit all tables to HTML. |
| `knitr::opts_chunk$set(fig.retina = 3)` | Makes images clearer.|


### bookdown

[*Slides*](https://arm.rbind.io/slides/bookdown.html)

In the console:

```{r, eval = FALSE}
# Start a bookdown from a template
bookdown::bookdown_skeleton(getwd())

# Render book
bookdown::render_book('index.Rmd')

# Serve book (powers "Preview book" add-in)
bookdown::serve_book()
```

Other tips:

- Cross-reference figures with `\@ref{fig:fig-chunk-name}` and tables with `\@ref{tab:tab-chunk-name}`
- It is easy to configure the appearence of the TOC Sidebar by adding options to the headers (see [slides 46-50](https://arm.rbind.io/slides/bookdown.html#46)) or configuring the `_output.yml` file ([slide 51](https://arm.rbind.io/slides/bookdown.html#51)).

In the YAML of the `index.Rmd`:

- For Github sharing, add `github-repo:mpaulacaldas/repo-name`.
- Use the `url`, `cover-image` and `description` options to customise your [card metadata](https://developer.twitter.com/en/docs/tweets/optimize-with-cards/overview/summary).

The `_bookdown.yml`:

- Gets fed to `bookdown::render_book()`.
- Output dir is set to `"docs"` to make publishing in Github pages easier.

The `_output.yml`: 

- Most arguments get passed to `rmarkdown::html_document()`.
- Don't forget to adapt the link in the `bookdown::gitbook:config:edit:link:` option so that it points to the correnct branch of your book's repo.

### blogdown

[*Slides*](https://arm.rbind.io/slides/blogdown.html)

Global configurations:

```{r, eval = FALSE}
library(usethis)

edit_r_profile()
options(
  blogdown.new_bundle = TRUE, # allows hugo page bundles
  blogdown.author = "Ada Lovelace",
  blogdown.ext = ".Rmd",
  blogdown.subdir = "blog",
  blogdown.yaml.empty = TRUE,
  blogdown.new_bundle = TRUE,
  blogdown.title_case = TRUE
  )
```

New site:

```{r, eval=FALSE}
library(blogdown)

new_site(
  theme = "jpescador/hugo-future-imperfect", 
  sample = TRUE, 
  theme_example = TRUE, 
  empty_dirs = TRUE,
  to_yaml = TRUE
  )
```
    
Keep in mind:

- Don't touch the `index.Rmd` or the `themes/` directory.
- Don't knit, serve site with `blogdown::serve_site()`.
- Ignore `public/`. Once you serve your site, you can drag this folder to Netlify.

Edit `config.toml`:

- `ignoreFiles`: Copy and paste from R console upon first serve of site.
- `relativeURLS = true` for cloud setup.
- `enableEmoji = true` obviously.
- `[params]` for more.

Workflow:

1. Open project
1. "Serve site"
1. "New post"
1. Write, save.
1. Commit, push.
1. On Netlify, *New site from Git > Pick site repo*

### Customization and extensions

[*Slides*](https://slides.yihui.name/2019-rstudio-conf-rmarkdown-workshop.html#1)

WIP

# Tidy eval

## Tidy eval in context, Jenny Brian

[Repo](https://github.com/jennybc/tidy-eval-context#readme), [Video](https://resources.rstudio.com/rstudio-conf-2019/lazy-evaluation)

You may **not** need tidy eval. Most people don't need to make direct use of [rlang](https://rlang.r-lib.org/).

| You want to ... | You need to know this much tidy eval |
|-----------------|--------------------------------------|
| Write simple functions to reduce duplication    | Pass the dots, `enquo()`, `!!` |
| Write functions that make names from user input | `:=` |
| Compute on expressions and manipulate environments | Know the theory + rlang | 

## Working with names and expressions in your tidy eval code, Lionel Henry

[Slides](https://speakerdeck.com/lionelhenry/selecting-and-doing-with-tidy-eval), [Video](https://resources.rstudio.com/rstudio-conf-2019/working-with-names-and-expressions-in-your-tidy-eval-code)

#### Context

There are two dplyr flavours:

- Action verbs: `mutate()`, `transmute()`, `summarise()`, `group_by()`
- Selection verbs: `select()`

Use scoped dplyr verbs (e.g. `group_by_at()`) to pass selections to action verbs.

#### "Passing the dots" strategies

1. Selections with dplyr
  
    ```{r}
    summary_functions <- list(
      ~ mean(., na.rm = TRUE),
      ~ sd(., na.rm = TRUE)
    )
    
    # Pass the dots!
    summarise_sels <- function(.data, ...) {
      summarise_at(.data, vars(...), summary_functions)
    }
    
    
    starwars %>% 
      summarise_sels(height:mass)
    ```

1. Actions with dplyr

    ```{r}
    # Pass the dots! 
    summarise_act <- function(.data, ...) {
      .data %>% 
        transmute(...) %>% 
        summarise_all(summary_functions)
    }
    
    starwars %>% 
      summarise_act(
        heightm = height / 100,
        bmi = mass / heightm^2 
      )
    ```

1. Gather with tidyr

    ```{r}
    gather_summarise_act <- function(.data, ...) {
      .data %>% 
        transmute(...) %>% 
        gather("Variable", "Value", -one_of(group_vars(.))) %>% 
        summarise_at(vars(Value), summary_functions)
    }
    
    starwars %>% 
      group_by(gender) %>% 
      gather_summarise_act(
        heightm = height / 100,
        bmi = mass / heightm^2 
      )
    ```
    
    
#### Takeaways

- "Passing the dots" strategies with dplyr (selection _and_ actions) work easily with grouped data.

- For other functions, like `tidyr::gather()`, we need to deal with the grouping variables inside the function.

- New discovery: [`dplyr::group_vars()`](https://dplyr.tidyverse.org/reference/groups.html)

# Other

## vctrs, Hadley Wickham

[Video](https://resources.rstudio.com/rstudio-conf-2019/vctrs-tools-for-making-size-and-type-consistent-functions)

- Package aimed at package developers.
- Base R leads to some **WAT** moments.
    ```{r}
    # Dates and date-times give unexpected output and are not consistent.
    
    today <- as.Date("2018-09-18")
    lunch <- as.POSIXct("2018-09-18 12:00", tz = "Europe/London")
    
    c(today, lunch)
    c(lunch, today)
    c(lunch)
    c(NULL, lunch)
    ```

- The `vctr` package tries to get rid of them.
    ```{r, error=TRUE}
    library(vctrs)
    
    # Forces you to be explicit about the type of output you want
    vec_c(1.5, "x")
    vec_c(1.5, "x", .ptype = "character")
    
    # Nonetheless, pragmatics > philosophy
    fa <- factor("a")
    fb <- factor("b")
    
    vec_c(fa, fb)
    vec_c(fb, fa)
    
    # No unexpected output, like in base R
    c(today, fa)
    vec_c(today, fa)
    ```
    
- And makes it easier to create new vector classes.

The **main principles** of the package:

1. The output type only depends on the types of the arguments.

    ```{r, results="hold"}
    # It's hard to guess the type of output returned by base::ifelse()
    typeof(ifelse(NA, 1, "a"))
    typeof(ifelse(TRUE, 1, "a"))
    typeof(ifelse(FALSE, 1, "a"))
    ```
    
    ```{r, error=TRUE}
    # dplyr is equally annoying
    typeof(if_else(NA, 1, "a"))
    ```

1. The order of inputs in `...` does not affect the output type.

1. It uses consistent rules.
    ```{r}
    # Needlessly exposes data scientist to R's internals
    c(factor("a"), factor("b"))
    
    # But doesn't follow an underlying principle
    unlist(list(factor("a"), factor("b")))
    ```





